---
title: "Peer Assessment 1"
output: html_document
---

**Loading and Processing the Data**

1. Load the Data 
```{r echo=TRUE}
activity<-read.csv("activity.csv")
```

2. Transform the data as needed
  + Updating Date to Date format 
  + Removing NAs
  + Placing in tbl format to use dpylr 
```{r echo=TRUE, warning=FALSE, message=FALSE}
activity$date<-as.Date(activity$date)
activity<-na.omit(activity)
library(dplyr)
activity<-tbl_df(activity)
```

**What is mean total number of steps taken per day?**

1. Calculate the total number of steps taken per day
```{r echo=TRUE}
daily_steps <- activity %>%
                group_by(date) %>%
                summarize(total_steps=sum(steps))
```

2. Make a histogram of Steps per Day
```{r echo=TRUE}
hist(daily_steps$total_steps, main="Total Steps per Day", xlab="Steps",col="purple")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
step_stats <- daily_steps %>%
                summarize(mean_steps=mean(total_steps),median_steps=median(total_steps))

```

- The Mean number of steps taken per day is equal to `r step_stats$mean_steps`  
- The Median number of steps taken per day is equal to `r step_stats$median_steps`  
  
**What is the average daily activity pattern?**

1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
  + First Summarize   
  + Then Plot   
```{r}
interval<- activity %>%
            group_by(interval) %>%
            summarize(mean_steps=mean(steps))

plot(interval$interval,interval$mean_steps,type="l",xlab="Interval",ylab = "Average Steps Taken",
     main="Average Steps taken during a 5 minute interval")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_int<-interval$interval[which.max(interval$mean_steps)]
```
- The Maximum average steps took place during interval `r max_int`
  
**Imputing missing values**  
1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)  
  + Since I originally removed the missings, I am reading back in the data  
  + Then using summary to see which variables have missing values  
```{r echo=TRUE}
activity2<-read.csv("activity.csv")
activity2$date<-as.Date(activity2$date)
summary(activity2)
```
- Since Steps is the only variable with missing values, that is the only variable I need to count for missings 
    
```{r echo=TRUE}
missings<-sum(is.na(activity2$steps))

```
- The total number of missing values is `r missings`

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.  
  + I am going to fill in the median number of average steps per interval step  
  + Average number of steps per interval is saved in interval already, just need to find median   
  
3.Create a new dataset that is equal to the original dataset but with the missing data filled in.  
```{r echo=TRUE}
int_med<- interval %>%
          summarize(med=median(mean_steps))

for (i in 1:nrow(activity2)) {
    if (is.na(activity2$steps[i])) {
        activity2$steps[i] <- int_med$med
    }
}
summary(activity2)
```
- There are no longer any missings   

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.  
  + First I will need to get total steps per day, then I can plot the histogram  
```{r echo=TRUE}
new_sum<- activity2 %>% 
          group_by(date)%>%
          summarize(total_steps=sum(steps))
hist(new_sum$total_steps, main="Total Steps per Day", xlab="Steps",col="purple")
```

- Mean and Median Steps   
```{r echo=TRUE}
new_stats<- new_sum %>%
            summarize(mean_steps=mean(total_steps),med_steps=mean(total_steps))
```

- These values are different:  
    - Mean  
        - Old `r step_stats$mean_steps`  
        - New `r new_stats$mean_steps`  
    - Median  
        - Old `r step_stats$median_steps`  
        - New `r new_stats$med_steps`
- The impact is mean average and median steps 

**Are there differences in activity patterns between weekdays and weekends?**
1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
  + Figure out Day of week with weekdays()
  + Assign daytype orginally to weekday
  + change Day Type on Sat/Sun
```{r echo=TRUE}
activity2$day<-as.factor(weekdays(activity2$date))
activity2$daytype<-"weekday"
activity2$daytype[activity2$day %in% c("Saturday", "Sunday")] <- "weekend"
activity2$daytype<-as.factor(activity2$daytype)
```

2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r echo=TRUE}
activity3 <- activity2 %>%
            group_by(daytype,interval) %>%
            summarize(mean_int=mean(steps))
library(ggplot2)
ggplot(activity3, aes(interval,mean_int)) +
                 ggtitle("Average Steps per Interval Steps Weekday vs Weekend") +
                 facet_grid(. ~ daytype) +
                 geom_line(size = 1, color="purple")
              


```
    

    